---
title: "Untitled"
format: html
editor: visual
---

```{r}
library(tidyverse)
```

## One single fit

```{r}
dat <- tibble(x = seq(0, 1, .05)) |> 
  rowwise() |> 
  mutate(y = rnorm(1, mean = 3 * x, sd = 1))

ggplot() +
  geom_point(data = dat, aes(x = x, y = y))
```

I would like to fit a line described by parameter p1 and p2. To do that I create a least square function that given some data points and some line specified by p1 and p2, creates a sum of the distances of the points to the line (square to make everything positive).

```{r}
least_square_fun <- function(.par) {
  y_hat <- .par[1] + .par[2] * dat$x 
  
  differences <- dat$y - y_hat 
  
  sum(differences^2)
  
}
```

where y hat is the value of y on the line for a given x.


I can compute the value of this function for different parameters. For example for a line with intercept 1 and slope 3:

```{r}
least_square_fun( c(1, 3))
```

For a line with intercept 0 and slope 4:

```{r}
least_square_fun( c(0, 4))
```

This last line is a better fit because the sum of least squares is better. Instead of trying by hand many parameters, we search for the parameters that minimize this function using the R function for optimization `optim`. It requires some initial parameters, let´s use the last ones that we tried, for example 

```{r}
fit <- optim(c(0, 4), least_square_fun)

fit
```

We can plot the fit now

```{r}
line <- tibble(x = seq(0, 1, .001)) |> 
  mutate(y = fit$par[1] + fit$par[2] * x)

ggplot() +
  geom_point(data = dat, aes(x = x, y = y)) +
  geom_line(data = line, aes(x = x, y = y))
```

### This is ugly

The problem is that it is not good coding creating a function that uses variables that are not an input to the function. That is, the function `least_square_fun` is using dat$x and dat$y. An elegant solution is to use function factories. A function factories is a function that given some inputs does not return a number, but instead in return another function that depends on other parameters. Let's do it. 

```{r}
create_ls <- function(.predictor, .response){
  function(.par) { #it returns a function that depend on .par
    y_hat <- .par[1] + .par[2] * .predictor 
    differences <- .response - y_hat 
    sum(differences^2)
  }
}
```


The function `create_ls` creates functions given some data. Let's create a function that uses the data from our experiment

```{r}
least_square_fun_for_my_exp <- create_ls(dat$x, dat$y)
```

The function `least_square_fun_for_my_exp` only depends on parameters. Let's compute its value for the values 0 and 4

```{r}
least_square_fun_for_my_exp(c(0, 4))
```

We can optimize now

```{r}
optim(c(0, 4), least_square_fun_for_my_exp)
```

So basically `least_square_fun_for_my_exp` is the same that our previous `least_square_fun` but now instead of using dat$x and dat$y inside the function we have a function factory function `create_ls` that introduce the data in a more elegant way. Given some other experiment, we can use `create_ls` to create our custome `least_square_fun` 

Now, we can be even more elegant by using an argument that takes into account the function that we want to fit

```{r}
create_ls <- function(.predictor, .response, .fun){
  function(.par) { 
    y_hat <- .fun(.predictor, .par)
    differences <- .response - y_hat 
    sum(differences^2)
  }
}
```


Let's use it for example to fit a linear model:

```{r}
linear_fun <- function(x, p) {
  p[1] + p[2] * x
}

least_square_fun <- create_ls(dat$x, dat$y, linear_fun)

optim(c(0, 4), least_square_fun)
```
The function `fit_ls` inside the folder R is an extension that also includes some initial parameters.

```{r}
fit <- fit_ls(dat$x, dat$y, linear_fun, c(0, 4))

fit
```

```{r}
ggplot() +
  geom_point(data = dat, aes(x = x, y = y)) +
  geom_line(data = fit$pred, aes(x = x, y = y))
```

## Multiple groups

## Creating some fake data


```{r}
dat_groups <- tibble(x = seq(0, 1, .1)) |> 
  rowwise() |> 
  mutate(y = rnorm(1, mean = 3 * x, sd = 2)) |> 
  mutate(group = "a") |> 
  bind_rows(
    tibble(x = seq(0, 1, .1)) |> 
      rowwise() |> 
      mutate(y = rnorm(1, mean = 5 * x, sd = 1)) |> 
      mutate(group = "b") 
  )
```

## Fit

```{r}
fit_groups <- dat_groups |> 
  group_by(group) |> 
  reframe(fit = list(fit_ls(x, y, linear_fun, c(0, 1)))) |> 
  unnest_wider(fit)
```

```{r}
params <- fit_groups |> 
  select(group, param) |> 
  unnest(param)

params
```

To get the fits 

```{r}
pred_groups <- fit_groups |> 
  select(group, pred) |> 
  unnest(pred)
```

```{r}
ggplot() +
  facet_wrap(vars(group)) +
  geom_point(data = dat_groups, aes(x = x, y = y)) +
  geom_line(data = pred_groups, aes(x = x, y = y))
```
