---
title: "Untitled"
format: html
editor: visual
---

This code reproduces some of the analysis in

Keysers, C., Gazzola, V., & Wagenmakers, E. J. (2020). Using Bayes factor hypothesis testing in neuroscience to establish evidence of absence. *Nature neuroscience*, *23*(7), 788-799.

```{r}
library(tidyverse)
library(broom)
library(BayesFactor)

list.files("R", full.names = TRUE) |> walk(source)
```


```{r}
dat <- read_csv("data/Muscimol1.csv", 
                show_col_types = FALSE)
```

## Experimental condition

In the experimental condition, it is hypothesized that muscimol changes freezing

```{r}
dat_ShockObs <- dat |> 
  select(Group, ShockObs)
```

```{r}
m_ci_ShockObs <- dat_ShockObs |> 
  group_by(Group) |> 
  summarise(calculate_t_ci(ShockObs))
```


```{r}
ggplot() +
  geom_point(data = dat_ShockObs , aes(x = Group, y = ShockObs)) +
  geom_crossbar(data = m_ci_ShockObs, 
                aes(x = Group, 
                    y = estimate, 
                    ymin = conf.low, 
                    ymax = conf.high))
```

We can perform a t-test

```{r}
t.test(ShockObs ~ Group, 
       data = dat_ShockObs)
```
There is a very significant effect of muscimol. 


Using Bayes factors

```{r}
ttestBF(formula = ShockObs ~ Group, 
        data = dat_ShockObs)
```
The hypothesis that freezing changes with muscimol is 81 times more likely than the hypothesis that freezing does not change with muscimol. 


## Control condition

```{r}
dat_CS <- dat |> 
  select(Group, CS)
```

```{r}
m_ci_CS <- dat_CS |> 
  group_by(Group) |> 
  summarise(calculate_t_ci(CS))
```

```{r}
ggplot() +
  geom_point(data = dat_CS , aes(x = Group, y = CS)) +
  geom_crossbar(data = m_ci_CS, 
                aes(x = Group, 
                    y = estimate, 
                    ymin = conf.low, 
                    ymax = conf.high))
```

```{r}
t.test(CS ~ Group, 
       data = dat_CS)
```
For the control, the effect is not significant. Why? There are two possibilities:

1) Muscimol does not affect freezing

2) We don't have enough data. 

We can use Bayes factors to differenciate between these two posssibilities. 


```{r}
ttestBF(formula = CS ~ Group, 
        data = dat_CS)
```
The hypothesis that freezing changes with muscimol is about 0.34 times more likely than the hypothesis that freezing does not change with muscimol.

This is equivalent to say that the hypothesis that freezing does not change with muscimol is 2.9 more likley that the hypothesis that freezing changes with muscimol. 

The reason is the mathematical definition of bayes factor: 


$$BF_{10} = \frac{P(\text{data} | H_1)}{P(\text{data} | H_0)} = 0.34$$

So 

$$BF_{01} = \frac{P(\text{data} | H_0)}{P(\text{data} | H_1)} = 1 /0.34 = 2.9$$




A rule of thumb (see the paper) to assess the eviden in favor of each hypothesis is: 

If BF_10 > 10 -> Strong support for H1

If 3 < BF_10 < 10 -> Moderate support for H1

If 1/3 < BF_10 < 3 -> More data in needed. There is not evidence for H0 and H1.

If 1/10 < BF_10 < 1/3 -> Moderate support for H0

If BF_10 < 1/10 -> Strong support for H0 


So, for the control condition we don't have enough evidence for the support of H0 or H1. 



## Directional test

For performing bayes factors, when possible, directional tests are recommended

t.test

```{r}
t.test(ShockObs ~ Group, 
       alternative = "less",
       data = dat_ShockObs)
```

Bayes factors

```{r}
ttestBF(formula = ShockObs ~ Group, 
        nullInterval = c(-Inf, 0), # This is the way to specify the directional hypothesis
        data = dat_ShockObs)
```
In the experimental condition, the hypothesis that freezing decreases with muscimol is 162 more likely than the hypothesis that muscimol does not affect freezing.

d is related to the difference between means (mu1 and m2). To know the corresponde betwen the mu1 and m2 and the levels, it is possible using:

```{r}
fct_unique(dat_ShockObs$Group)
```



For the control:

t.test

```{r}
t.test(CS ~ Group, 
       alternative = "less",
       data = dat_CS)
```

Bayes factors

```{r}
ttestBF(formula = CS ~ Group, 
        nullInterval = c(-Inf, 0), # This is the way to specify the directional hypothesis
        data = dat_CS)
```
So BF_10 = 0.222

which means that we have moderate evidence for the support of H0.


Notice that BF_01 = 1 / 0.222 = 4.5 which means that no effect is 4.5 times more likely than the hypothesis by which muscimos reduces freezing. 


